---
title: "Assignment week 5"
author: "Sarah Eiben"
date: '2022-03-30'
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(Rsubread)
  library(rtracklayer)
  library(Biostrings)
  library(Rfastp)
  library(epiwraps)
})
ah <- AnnotationHub()
ensdb <- ah[["AH89211"]]
```

# Install the new version of epiwraps

```{r}
BiocManager::install("ETHZ-INS/epiwraps")
```

# download peaks and tracks for H3K4me3 and H3K27me3

```{r, eval=FALSE}
dir.create("peaks")
dir.create("tracks")
#change timeout for internet operations from 60s to 1800s
options(timeout=1800)

# H3K27me3
download.file("https://www.encodeproject.org/files/ENCFF008XKX/@@download/ENCFF008XKX.bed.gz", "peaks/H3K27me3.bed.gz")
download.file("https://www.encodeproject.org/files/ENCFF160FEV/@@download/ENCFF160FEV.bigWig", "tracks/H3K27me3.bw")

# H3K4me3
download.file("https://www.encodeproject.org/files/ENCFF247GVM/@@download/ENCFF247GVM.bed.gz", "peaks/H3K4me3.bed.gz")
download.file("https://www.encodeproject.org/files/ENCFF611GSQ/@@download/ENCFF611GSQ.bigWig", "tracks/H3K4me3.bw")
```
# Reading in the data

```{r}
k4me3 <- import("peaks/H3K4me3.bed.gz", format="NarrowPeak")
k27me3 <- import("peaks/H3K27me3.bed.gz", format="NarrowPeak")
```

# Identify bivalent domains

Bivalent domains are regions where both H3K4me3 AND H3K27me3 bind and let it stay in standby.

## Find overlaps between k4me3 and k27me3

```{r}
#overlapsAny gives you True and False, sum adds up how many there are
sum(overlapsAny(k4me3, k27me3))
```
There are 3219 Overlaps between k4me3 and k27me3, which means there are 3219 bivalent regions.


```{r}
listofRegions <- list(k4me3=k4me3, k27me3=k27me3)
#here the regionOverlaps for some reason doesn't work? Do there have to be more than 2 inputs?
#regionOverlaps(listofRegions)
regionUpset(listofRegions)
```
Shows 2746 Overlaps.

# Download smooth muscle cell peaks (h3k27me3, h3k4me3)

```{r, eval=FALSE}
download.file("https://www.encodeproject.org/files/ENCFF369AEO/@@download/ENCFF369AEO.bed.gz", "peaks/smoothH3K27me3.bed.gz")

download.file("https://www.encodeproject.org/files/ENCFF965WTD/@@download/ENCFF965WTD.bed.gz", "peaks/smoothH3K4me3.bed.gz")
```
# Reading in the data

```{r}
smoothk27me3 <- import("peaks/smoothH3K27me3.bed.gz", format = "NarrowPeak")
smoothk4me3 <- import("peaks/smoothH3K4me3.bed.gz", format = "NarrowPeak")
```

# Overlaps of the mESC with the smooth muscle cell peaks

## Overlaps of the k27me3 
```{r}
sum(overlapsAny(k27me3, smoothk27me3))
```
Overlaps: 232

## Overlaps of the k4me3 
```{r}
sum(overlapsAny(k4me3, smoothk4me3))
```
Overlaps: 1542

## Overlaps of all of them
```{r}
smoothListofRegions <- list(k4me3=k4me3, k27me3=k27me3, smoothk27me3=smoothk27me3, smoothk4me3=smoothk4me3)

#Create a heatmap for this
#Question: why isn't the yellow and the orange field the same?
regionOverlaps(smoothListofRegions)
#Create a histogram for this
#Question: why isn't this number for the overlaps of k27me3 and k4me3 the same as when doing a regionUpset with only those 2?
regionUpset(smoothListofRegions)
```
Estimated time for completion: 2h