---
title: "Assignment Week 6"
author: "Sarah Eiben"
date: '2022-04-06'
output: html_document
---

Question: How do I pick one peak file from encode? I sometimes struggle with finding an optimal one.
What exactly does the wrapper function do?


```{r}
suppressPackageStartupMessages({
  library(GenomicRanges)
  library(memes)
  library(Biostrings)
  library(AnnotationHub)
  library(MotifDb)
  library(universalmotif)
  library(PWMEnrich)
  library(ggplot2)
})

ah <- AnnotationHub()
```

# Choosing the transcription factor GATA1 in mice and downloading the peaks from the G1E cell line

We are also preparing the peaks for usage.

```{r}
dir.create("peaks")
download.file("https://www.encodeproject.org/files/ENCFF045RXZ/@@download/ENCFF045RXZ.bed.gz", "peaks/GATA1_G1E.bed.gz")
peaks <- rtracklayer::import("peaks/GATA1_G1E.bed.gz", format = "NarrowPeak")
seqlevelsStyle(peaks) <- "Ensembl"
```

## We are using the centers of the peaks so it doesn't run as long

```{r}
peak_centers <- resize(peaks, fix="center", width=100)
# Two bit file of Genome sequence of Mus Musculus GRCm38
# needs to be two bit file for meme
genome <- ah[["AH88477"]]
peak_seqs <- memes::get_sequence(peak_centers, genome)
peak_seqs
# see if there is a difference if we use the whole peaks
# peak_seqs_full <- memes::get_sequence(peaks, genome)
# peak_seqs_full
# here we get the same result as if we use only the peak centers.
```

# Motif scanning

Motif scanning aims at finding the _occurrences_ of a known motif in a set of sequences.

## Getting the desired motif (in our case this is GATA1)

```{r}
# we search for "GATA1" in the motif database
motifs <- query(MotifDb, "GATA1")
# there are several (to be exact 31) matching motifs (that just match the word GATA1):
names(motifs)
# we select one from the list of names (JASPAR is good, second best is HOCOMOCO):
# motif_2 <- motifs[["Mmusculus-jaspar2018-Gata1-MA0035.1"]]
# we visualize it (this is the motif of GATA1):
# view_motifs(motif_2)
# trying a different one just to see
motif <- motifs[["Mmusculus-jaspar2018-Tal1::Gata1-MA0140.1"]]
view_motifs(motif)
# I like this one better so I will continue with this one.
```

# Scanning a set of sequences with the slower alternative because I'm using Windows

## first the wrapper function

```{r}
findMotifInstances <- function(seqs, motif, keepMatchedSeq=FALSE, ...){
  motif <- universalmotif::convert_motifs(motif, "TFBSTools-PWMatrix")
  x <- suppressWarnings(TFBSTools::searchSeq(motif, subject=seqs))
  peaks <- strsplit(gsub("-",":",names(seqs)), ":")
  if(all(lengths(peaks)==3)){ # convert relative coordinates to absolute
    chrs <- sapply(peaks,FUN=function(x) x[1])
    offsets <- sapply(peaks,FUN=function(x) as.integer(x[2]))-1L
    i <- rep(seq_along(x),lengths(x))
    return(GRanges(chrs[i], 
                   IRanges(offsets[i]+as.integer(unlist(lapply(x, start))),
                           offsets[i]+as.integer(unlist(lapply(x, end)))),
                   strand=unlist(lapply(x, strand)),
                   score=as.numeric(unlist(lapply(x, FUN=function(x) score(x))))
                   ))
  }
  x <- as(x, "GRanges")
  score(x) <- x$absScore
  keepFields <- c("score","relScore")
  if(keepMatchedSeq) keepFields <- c(keepFields, "siteSeqs")
  mcols(x) <- mcols(x)[,intersect(colnames(mcols(x)), keepFields)]
  x
}
```

## Then get the motif instances

```{r}
motif_instances <- findMotifInstances(peak_seqs, motif, mc.cores=2) # running with 2 threads
```

# Answering the questions
## Of all the peaks, what proportion contains a motif for the factor?

```{r}
# all the peaks:
length(peaks)
# peaks with motif
length(motif_instances)
# why does the following give me a smaller number than length(motif_instances)? shouldn't it be the same?
sum(overlapsAny(peaks, motif_instances))
percentage <- (sum(overlapsAny(peaks, motif_instances)))/length(peaks)*100
percentage
```
Of  the `r length(peaks)` peaks, `r sum(overlapsAny(peaks, motif_instances))` (`r percentage`%) contain a motif.

## Of all instances of that motif in the genome, what proportion is bound by the factor (i.e. has a peak)?

```{r}
mmusculus <- import(genome, "2bit", which = as(seqinfo(genome), "GenomicRanges"))
motif_instances_genome <- findMotifInstances(mmusculus, motif, mc.cores=2)
length(motif_instances_genome)
```

```{r}
motif_with_peaks = overlapsAny(motif_instances_genome, peaks)
sum(motif_with_peaks)
percentage2 <- sum(motif_with_peaks)/length(motif_instances_genome)*100
percentage2
```

Of the `r length(motif_instances_genome)` motif instances, `r sum(motif_with_peaks)` (`r percentage2`%) overlap a peak.
